---
title: ''
output:
  html_document:
    css:
    - style.css
    - https://maxcdn.bootstrapcdn.com/bootstrap/4.0.0/css/bootstrap.min.css
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r DATAIMPORT, echo = FALSE, message = FALSE, warning= FALSE}
rm(list = ls())
library(tidyverse)
library(leaflet)
library(dplyr)
library(readr)
library(jsonlite)

# Data on Drug accidents in the US 2015-2018
drug_set <- read_csv("https://data.cdc.gov/api/views/xkb8-kh2a/rows.csv?accessType=DOWNLOAD")

# Adding states names to dataset

drug_set <- drug_set %>% mutate(state = case_when(State == "AK" ~ "Alaska",
                                                  State == "AL" ~ "Alabama",
                                                  State == "AZ" ~ "Arizona",
                                                  State == "AR" ~ "Arkansas",
                                                  State == "CA" ~ "California",
                                                  State == "CO" ~ "Colorado",
                                                  State == "CT" ~ "Connecticut",
                                                  State == "DC" ~ "District of Columbia",
                                                  State == "DE" ~ "Delaware",
                                                  State == "FL" ~ "Florida",
                                                  State == "GA" ~ "Georgia",
                                                  State == "HI" ~ "Hawaii",
                                                  State == "ID" ~ "Idaho",
                                                  State == "IL" ~ "Illinois",
                                                  State == "IN" ~ "Indiana",
                                                  State == "IA" ~ "Iowa",
                                                  State == "KS" ~ "Kansas",
                                                  State == "KY" ~ "Kentucky",
                                                  State == "LA" ~ "Louisiana",
                                                  State == "ME" ~ "Maine",
                                                  State == "MD" ~ "Maryland",
                                                  State == "MA" ~ "Massachusetts",
                                                  State == "MI" ~ "Michigan",
                                                  State == "MN" ~ "Minnesota",
                                                  State == "MS" ~ "Mississippi",
                                                  State == "MO" ~ "Missouri",
                                                  State == "MT" ~ "Montana",
                                                  State == "NE" ~ "Nebraska",
                                                  State == "NV" ~ "Nevada",
                                                  State == "NH" ~ "New Hampshire",
                                                  State == "NJ" ~ "New Jersey",
                                                  State == "NM" ~ "New Mexico",
                                                  State == "NY" ~ "New York",
                                                  State == "NC" ~ "North Carolina",
                                                  State == "ND" ~ "North Dakota",
                                                  State == "OH" ~ "Ohio",
                                                  State == "OK" ~ "Oklahoma",
                                                  State == "OR" ~ "Oregon",
                                                  State == "PA" ~ "Pennsylvania",
                                                  State == "RI" ~ "Rhode Island",
                                                  State == "SC" ~ "South Carolina",
                                                  State == "SD" ~ "South Dakota",
                                                  State == "TN" ~ "Tennessee",
                                                  State == "TX" ~ "Texas",
                                                  State == "US" ~ "United States of America federal state",
                                                  State == "UT" ~ "Utah",
                                                  State == "VT" ~ "Vermont",
                                                  State == "VA" ~ "Virginia", 
                                                  State == "WA" ~ "Washington",
                                                  State == "WV" ~ "West Virginia",
                                                  State == "WI" ~ "Wisconsin",
                                                  State == "WY" ~ "Wyoming",
                                                  State == "YC" ~ "Yucatan(MEX)"))
# Changing colnames 
colnames(drug_set) <- c("State","State_abbreviations","Year", "Month", "Indicator", "Deaths", "Predicted_Value", "Percent_complete", "Percent_pending_investigation","Footnote","State_name")

# Removing the first colum we have two of                                                                       
drug_set$State <- NULL

# Rearranging colums 
drug_set <- drug_set[,c(10,1,2,3,4,5,6,7,8,9)]


# Taking a look at the dataset after cleaning. 
head(drug_set)

# Examining the footnote colum
unique(drug_set$Footnote)
unique(drug_set$Year)

# If the code below is runned the data for 2017 and 2018 are removed. This is because the investigation on cases not are 
# juridicaly finished and solved and some cases are underreported due to incomplete data. We still want to use the data for 2017 and 2018.

# drug_set <- drug_set %>%
  # filter(is.na(Footnote))




# Data on medianincome from api
getdata <- fromJSON("https://api.datausa.io/api/?sort=desc&show=geo&required=income&sumlevel=state&year=all&where=geo%3A%5E04000US")

#Getting data out of list
income <- getdata[["data"]]

#Labeling data 
colnames(income)[1:3] <- c("year","state", "median_income")

#Exporting data to a csv file to avoid problems with atomic vectors
write.csv(income, file = "MyData.csv")

#Importing the csv
income <- read_csv("mydata.csv", col_types = cols(X1 = col_skip()))


```




```{r}
unique(drug_set$Indicator)


```



<!--Jumbotron -->



<div class="jumbotron jumbotron-fluid" id="background_div">
<h1 class = "text-center" >Drug related deaths in Conneticut</h1>




<!-- Introduction text -->
<div class = "container" id  = "introduction">
<hr style ="margin-bottom:10%;" />

  <h2 class = "text-center"> Introduction </h2>
  <p class = "text-center"> In this casestudy we're going to look at drug related death in Conneticut ipsoms lorums ripsbusker og andre buskvekster ipsoms lorums ripsbusker og andre buskvekster ipsoms lorums ripsbusker og andre buskvekster ipsoms lorums ripsbusker og andre buskvekster  </p> 
  
<hr style = margin-top:10% />




<!--Numbers -->
<div id="numbers">   
<div class="container-fluid" id="numbers_container">   

<div class="card">
<center>
```{r echo = FALSE, message= FALSE}
# ggplot() +
#   geom_bar(data = drugsidents, aes (Race)) +
#   labs(
#         title = "Races",
#         y = "Count") +
#     theme_dark()
```
</center>

<div class="card-body">
<p class="card-text text-center">Drug related use up by going to look at drug related death in Conneticut ipsoms lorums ripsbusker og andre buskvekster ipsoms lorums ripsbuske </p>
</div>
</div>




</div>
</div>



<!--Maps -->
<div class = "container" id  = "maps">
  <h2 class = "text-center"> Map </h2>

<center>
```{r map, echo = FALSE, message= FALSE, fig.width=8}
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
m
```
</center>





<!--Statistics -->


<div id ="stat">
<h2 class = "text-center"> Statistics </h2>

<div class="container-fluid" id="stat_container">   

<div class="col-md-6" id = "graph1">
```{r graph, echo = FALSE, message= FALSE}
# ggplot() +
#   geom_bar(data = drugsidents, aes (Race)) +
#   labs(
#         title = "Races",
#         y = "Count") 

```
</div>


<div class="col-md-6" id = "textgraph">
<p> ipsoms lorums ripsbusker og andre buskvekster ipsoms lorums ripsbusker og andre buskvekster ipsoms lorums ripsbusker og andre buskvekster </p>
</div>


</div>
</div>









</div>


  






